Hysteresis in human computation:\\ how one task affects another

### 2014-12-03 docs/publications/2014.09_SCIENCE/man.tex
re-map figs symlink
substancially shorten science manuscript

Microtask platforms combine the efforts of large numbers 
of people to perform tasks that are difficult to automate with computers 
alone.  These platforms support near-real-time task completion, mimicking a
compute server, and shift the role of the human from consumer to 
producer of computing resources.  To realize its full
potential, we must understand how human computation differs from machine 
computation, and address the biases inherent in 
human cognition.  The effects of various task-design variables on worker
performance have been investigated, but there has been little 
investigation into how consecutive tasks influence one another. Here we show 
that the responses in later tasks are strongly influenced by the content of 
earlier ones, an effect that is as strong or stronger than the effect of 
framing.  Our findings are not solely cautionary; we find
that these \textit{intertask effects} can be harnessed to improve the quality 
of responses, and might thereby help to achieve expert-level judgments in 
citizen science and other crowdsourcing initiatives. 
As an ancillary contribution, we demonstrate the use of machine learning 
algorithms measure exposure-induced human bias, in terms of its ability to 
influence human computations. 
\end{sciabstract}

\section*{Introduction}
Microtask crowdsourcing platforms like Amazon Mechanical Turk (MTurk) make it 
possible for requesters to submit batches of tasks to a large pool of 
workers, who do the tasks for fun, a sense of purpose, and remuneration 
\cite{kazai2013analysis,Antin20122925}.  
Typical tasks include tagging and categorizing images 
\cite{6116320,Zhai2012357}, transcribing voice recordings 
\cite{chandler2013breaking,paolacci2010running}
or handwritten notes \cite{Berinsky2012351,Finnerty2013}, and judging the 
relevancy of search results 
\cite{le2010ensuring,grady2010crowdsourcing,alonso2009can,kazai2013analysis}.
Originally used to distribute clerical work (\#), these platforms 
increasingly serve as a means to engage experimental participants in a 
research 
setting \cite{paolacci2010running,Berinsky2012351,snow2008cheap,alonso2009can}.
This trend has seeded considerable interest in characterizing the parameters
of task design, and validating the microtask platform methodology (\#).

Computer scientists view microtask platforms in the context of general
computation(\#).  In this regard, the platforms enable a tight integration of 
human and machine computation(\#).  Researchers coined the term HPU 
(Human co-Processing Unit) in analogy to CPUs and GPUs \cite{5543192}.  
Ongoing research pursues the establishment of an HPU instruction set
including basic operators like \texttt{vote}, \texttt{subdivide}, \texttt{map},
\texttt{reduce}, \texttt{sentiment}, and \texttt{image\_tag} (\#).  Libraries 
are being developed to give the programmer more seamless access to HPU 
resources 
 \cite{little2010turkit,minder2011crowdlang,minder2012crowdlang,kittur2011crowdforge}.  
As with other architectures, the HPU architecture makes entirely new classes 
of problems tractable for the programmer (\#).

But the ability to program with high-level human instructions comes with a 
cost: HPU output is inherently noisy (\#).  Part of this noise can be explained
in terms of worker demography (\#).  But systematic cognitive biases also 
play a role (\#).   
So far, investigation has explored how worker performance is influenced by
such design factors as the level of 
pay \cite{kazai2013analysis}, training \cite{le2010ensuring}, pre-screening of 
workers \cite{paolacci2010running}, and user-interface design 
\cite{Finnerty2013}.  Researchers have also investigated \textit{framing}, 
by testing the effects of describing the workflow context 
\cite{Kinnaird2012281}, the purpose of tasks 
\cite{chandler2013breaking}, or using alternate problem descriptions
\cite{thibodeau2013natural}.  To our knowledge, no study has investigated 
the influences that consecutive microtasks exert on one another.

It is known that people are susceptible to priming effects 
\cite{BJOP1796,No2007,beller1971priming}, and, in particular, task-repetition 
effects \cite{Gass1999549,sohn2001task}.  Thus, a worker's response during
one task may depend on previous ones.  Such \textit{intertask} effects
would amount to a kind of \textit{hysteresis}, meaning that HPU outputs are not
only a function of the current input, but also of the history of inputs, with
important implications for the design of crowdsourcing initiatives.

Here we describe a series of experiments, using the Amazon Mechanical
Turk (MTurk) microtask platform (\#), in which we studied intertask effects
for image-labeling tasks.  Image-labeling tasks are among the most 
common kinds of microtasks
\cite{chandler2013breaking,Berinsky2012351,Finnerty2013,paolacci2010running}, 
and it seems likely that this will remain the case in the long term because
they provide a source of ground truth in computer vision research 
\cite{5543192}.  
During the experiments, workers were presented with a series of images (one at
a time), and had to provide five descriptive words (\textit{labels}) for each.
We varied first five images (hereafter the \textit{initial tasks})
and observed the effect this had on the workers' labels for the last five
(\textit{test tasks}).

The initial tasks were chosen to embed a certain \textit{prime concept}.
Our first experiment compares the prime concepts \textit{food} and (non-food) 
\textit{objects}, while our second experiment compares \textit{food} and 
\textit{culture}.  For both experiments, The test tasks used images that
combined both prime concepts used for the initial tasks \td{reword}.  Workers participated 
in only one treatment of one experiment, and were randomly assigned initial 
tasks based on one of the prime concepts.  From the perspective of the worker, 
we made no distinction between the initial tasks and test tasks.

As a point of comparison, we included treatments in which we induced
\textit{framing}. \td{what is framing?}
In the framing treatments, before workers begin the tasks, we told them that 
the purpose of study was to understand the visual perception of 
``food and ingredients'' in one case, or the 
perception of ``objects and tools'' in another.  We varied the language of 
the framing text (hereafter simply the \textit{frame}), and for some treatments
we required the worker to echo the frame using a combo-box input.

We found that intertask effects were stronger than framing, unless the worker
was required to echo the frame, in which case the effects of both exposures were on par.  
Our results demonstrate that initial tasks can change 
workers' focus in later tasks.  Using the wordnet corpus, we analyzed 
the effects on the vocabulary used by workers when 
labeling test tasks. 
Exposure to both framing and initial tasks significantly altered the 
diversity of vocabulary that workers used when referring to the prime concept,
as well as the degree to which workers used specialized (rather than generic) 
words.

Contrary to our initial expectations, our findings suggest that the HPU 
programmer can leverage intertask effects.  By carefully designing workflows,
it is possible to tune worker focus, specificity, and expressivity.  Our 
findings carry two lessons for the task designer: intertask
effects can severely bias HPU output if uncontrolled, but if properly 
controlled, intertask effects might be of significant utility in achieving 
expert-level output.

\paragraph*{Measuring computational hysteresis.}
Before describing the details of our experiments, we explain some 
choices made in measuring intertask and framing effects 
(hereafter collectively \textit{exposure effects}).  
In trying to establish a measure of exposure effects, it might be tempting to 
invoke the notion of a ``neutral frame of mind'' to serve as a baseline. 
But such a concept seems both difficult to define, and unlikely to arise.
Therefore, we always consider the effect of one exposure 
\textit{relative} to another.  We organized our experiments 
into pairs of treatments, between which it is possible to make meaningful 
comparisons.

Our goal in measuring exposure effects is to show that they can actually 
alter the course of algorithms.
We could, by means of a $\chi^2$ test, reject the null hypothesis that
differently exposed HPUs have the same distribution of outputs.
However, this alone would not demonstrate a practically significant effect, in 
terms altering the execution path of algorithms.

We developed a more stringent measure in which we train classifier algorithms 
to infer the exposure of an HPU from its outputs.
If a classifier infers an HPU's exposure with significantly 
better accuracy than provided by chance, then the HPU outputs must have 
been affected by the exposure.  
But in addition, the classifier serves as a kind of certificate, since it is 
an example of an algorithm whose path of execution is influenced by the 
exposure of 
individual HPUs.  In the supplementary material, we ground our approach 
rigorously in terms of the L1-distance metric, which is a measure of 
divergence between distributions.  The L1-distance ranges from 
0 to 1.  In our case, a value of 1 indicates that the HPU's exposure
influences its output so strongly that the exposure can be inferred from the
output without ambiguity.  A value of 0 indicates that the exposure has no
impact on the HPU's outputs.  The connection between the
L1-distance, $D_\mathrm{L1}$, and a classifier's accuracy, $\eta$, is given by:
\begin{equation}
	D_\mathrm{L1} \geq 2 \eta - 1, 
	\label{l1}
\end{equation}
Since a classifier's performance provides a lower bound on the L1-distance, 
we shall denote $D_\mathrm{L1}^-\equiv 2\eta-1$.
The L1-distance admits the following interpretation: if two distributions
$P$ and $Q$ have an L1-distance $d$, then in the worst case, the 
frequency with which a binary algorithm operating on samples
from $P$ outputs a 1, will differ from that when operating on samples from 
$Q$ by $d \times 100\%$.  
Thus, we can interpret $D_\mathrm{L1}^-$ as a kind of
lower bound for the worst-case bias induced by HPU hysteresis.
We provide further discussion of the statistical properties of our approach 
in the supplementary material.

\paragraph{Experimental setup.}
\begin{figure}
	\centering
	\includegraphics[scale=0.7]{figs/images.pdf}
	\caption{
		Examples of images used in labeling tasks. (The italicized treatment 
		names that follow are defined in Table \ref{table:treatments}.) Images 
		from initial tasks: A) \textit{task1:food}, B) 
		\textit{task1:obj}, D) \textit{task2:food}, 
		E) \textit{task2:cult}.  Images from  test tasks: C) \textit{exp1}, 
		F) \textit{exp2}.  The full set of experimental materials is 
		shown in the supplementary material (S1).
	}

	\label{fig:task}
\end{figure}
We performed two experiments, soliciting respectively 2300, and 900 MTurk 
workers to label images depicting food, kitchen- and dining-ware, and cultural
scenes.  In a given experiment, all workers performed the same test
tasks, the output of which was our main dependent variable.  Before performing
the test tasks, workers were subjected to an exposure, via two 
\textit{modalities}: either they completed a set of initial tasks, or 
were exposed to a frame, or both.  As mentioned, from the perspective of the worker, 
there was no distinction or discontinuity between initial and test tasks.

We organized the experiments into \textit{treatment pairs}.  Both treatments in a 
pair involved the same exposure modality, but differed in the
content of the exposure.  In Table 1, we summarize the various experimental
treatments, and introduce a naming convention that we use to refer to 
treatments and treatment pairs.
In the treatment pairs that studied intertask effects (\textit{task1} and
\textit{task2}), workers performed 10 image-labeling tasks, with the first 
five tasks comprising the initial tasks and the last five tasks comprising the
test tasks.

In the first experiment, the content of the exposures was based on the prime
concepts \textit{food} and \textit{objects} (throughout, ``objects'' should be
taken to imply ``non-edible objects'').  So, 
in \textit{task1:food}, the initial tasks contained images depicting food, 
while in \textit{task1:obj}, they contained images depicting objects, 
including kitchen and dining-ware (see Fig~\ref{fig:task}).  Meanwhile, 
for the exposures in \textit{frame1} we indicating that the study was
``funded by the laboratory for the visual perception of \{ Food and Ingredients
$\vert$ Objects and Tools~\}''.  To further strengthen the frame, we 
introduced another pair of framing treatments, \textit{echo}, with stronger
wording: ``The purpose of this study is to understand the visual perception
of \{ Food and Ingredients $\vert$ Objects and Tools~\}'', and required
the worker to echo the purpose of the study by selecting it among other 
options in a combo-box input.

Note that food and objects are both relatively well-defined classes of 
physical entities.  To explore a different conceptual axis, in 
\textit{exp2} we adopted the prime concepts \textit{food} and \textit{culture}.
Compared to food and objects, culture is more abstract, and, rather than 
comprising a class of physical entities, it encapsulates
sets of beliefs, values, and practices.  Examples of images from the initial 
tasks in \textit{task2} are shown in Fig.~\ref{fig:task}, and the full 
collection of 
experimental materials is presented in the supplementary material (S1).  We 
ensured that all workers only participated in one treatment of one experiment.

@@

	\caption{ \footnotesize{ 
		We use the treatment names shown to refer to specific sections of 
		the data collected.  For example, \textit{task2:food} refers to the
		treatment in the second experiment which was not exposed to a frame, 
		and whose initial tasks contained images of food (7th row).
	}}
	\label{table:treatments}
\end{table}

\paragraph{The strength of intertask effects.}

We assessed the strength of intertask and framing effects
by training a multinomial naive Bayes classifier to infer a worker's exposure
based on the labels provided in the test tasks (see Fig.~\ref{fig:theta}).
This inference was made over the two possibilities for a given treatment pair.
We found that in \textit{task1}, whether a worker performed food- or 
object-containing initial tasks had a strong effect on their subsequent 
labels, introducing a bias that probably exceeds 30\%.  In comparison, the 
framing effect in \textit{frame1} did 
not influence workers' labels enough for the classifier to make predictions 
with significantly better accuracy than that afforded by chance.  The results 
for \textit{exp2} show a similar, but even more pronounced trend, with the bias
induced by intertask effects in \textit{task2} probably exceeding 50\%.

We see a comparable effect size in the \textit{echo} treatment pair.
We expected \textit{echo} to show a very strong effect because, by requiring
the worker to echo the purported purpose of the task, we signal that it
is of particular importance.  The fact that the intertask effects were on par with the framing 
effects in \textit{echo} shows just how powerful intertask effects can be.

@@

\paragraph{Dynamics of intertask effects.} It is natural to expect 
intertask effects to attenuate as a worker proceeds through the test tasks.  
In \textit{task1}, we permuted the test tasks, allowing us to 
measure intertask effects as a function of task position, while factoring
out the content-dependent differences between tasks
(see Fig.~\ref{fig:theta}B). 
We found intertask effects were most pronounced for the first test task,
dropping off substantially for the second, but maintaining a 
significant effect through until the fifth test task.  
This shows that the range of intertask effects extends across at least five 
tasks, and probably substantially more.
This makes intertask effects of greater concern, but as we will show, it also
makes intertask effects of greater potential utility.
In future work, it would be interesting to investigate whether a worker's
susceptibility to intertask effects varies across a large batch.  For example,
It is conceivable that the first few tasks induce especially strong
or long-lasting effects, in effect ``setting the tone'' for the rest of the batch.

@@

Having observed significant effects from framing and task-exposure,
we looked in greater detail at the nature of the effects.  For instance,
does a food-oriented exposure induce workers to provide more food-related
labels?  And if so, does this explain the classifier's ability to distinguish
a worker's initial exposure?

To answer these questions, we used the wordnet corpus, augmented with names of
ethnic foods learned from the Internet.  Wordnet provides a (roughly)
hierarchical set of relationships between nouns, called hypernym-hyponym
relationships.  A hypernym is a generalization, while a hyponym is a 
specialization.  So, for example, ``bread'' is hypernym to ``rye bread'' 
and hyponym to ``baked goods''.

Using the augmented hypernym-hyponym structure of wordnet to identify nouns 
depicting food, we found that, for the experimental pairings in 
\textit{exp1}, the
food-exposed treatments did produce a significantly larger fraction of 
food-related labels (see Fig.~\ref{fig:specificity}A).  While the shift in 
focus with regard to food does help explain the classifier's ability to 
infer worker exposures, it is insufficient to explain the observed performance
(*).
Furthermore, in \textit{exp2:task2} we see a significant but reversed effect
(see Fig.~\ref{fig:specificity}A), which suggests a more complicated 
relationship between exposure and label usage.
Given the difference between the food-culture and food-object conceptual axes, 
it should perhaps not be surprising that the effect of initial task exposure 
is different.  It would appear that we must be prepared to accommodate 
multiple 
countervailing factors influencing the number of references to the prime 
concept (food).  We will come back to this point momentarily.

To explore these effects further, we looked more closely at which particular
labels were most suppressed (activated) by a worker's exposure.  Remarkably,
we found that, for all treatment pairs, ``food'' was always the most 
suppressed label among 
the food-exposed workers.  This is surprising, considering 
that the overall incidence of food-related labels mostly \textit{increased} 
among food-exposed workers.  But taken together, the increase in references to 
food 
and the decrease in the incidence of ``food'' per se (which is the most generic
food-reference in the wordnet-based ontology), necessarily implies that
food-exposed workers have opted for more specialized references in place of 
``food''.

In a crowdsourcing task, considerable effort must be given to shaping the 
vocabulary with which a worker engages a task.  To draw an example from the 
domain of citizen science, in galaxy zoo, workers catalog images of galaxies 
in terms of \textit{smoothness}, and the presence of \textit{rings}, 
\textit{dust lanes}, \textit{digital disturbances}, and so on.  Shaping this 
vocabulary is crucial to the reliability of results.  The interface of
galaxy zoo helps by constraining and prompting the user with this basic 
vocabulary.  But this would not appear to be sufficient, since
a set of positive and negative training examples are used, together with
peer-to-peer and peer-to-expert exchange in discussion areas, presumably to
further refine and standardize the vocabulary (\#).

In our set-up, where the worker's vocabulary is unconstrained, we can 
directly observe the effects that frames and initial tasks have on worker 
vocabulary.  By looking at the number of unique food-related labels that 
workers produced, we assessed the richness of vocabulary that workers used when
referring to food (see Fig.~\ref{fig:specificity}B).  We found that the 
food-exposed workers use a richer vocabulary of 
food-related words, even in cases like \textit{task2:food}, where they 
made fewer references to food overall.  
Interestingly, task-based exposure to food had a stronger enriching effect 
 than did framing-based exposures. (* explain why vocab 
diversity is useful.)

Using the ontology of food induced by wordnet, along with foods learned from
the Internet, we were able to directly calculate the relative specificity of 
workers from different treatments.
To do this, we considered all the labels from one treatment
(for example \textit{task1:food}) and compared these to all the labels from 
another (\textit{task1:obj}).  For all possible comparisons, we observed 
which treatment's label was more specific more often, and normalized this
quantity as a percentage (we provide greater detail in the supplementary 
material).

We calculated the relative specificity for all treatment pairs, while 
restricting focus to 
food-related labels (see Fig.~\ref{fig:specificity}C).  We found that 
the food-exposed workers always provided food-references of substantially 
greater specificity.  Workers that had labeled images containing food 
were about 12\% more specific than those that labeled images of objects,
and about 20\% more specific than those that labeled images of cultural 
scenes.  With respect to the specificity of food references, framing effects 
were on par with intertask effects.

Taken together, the results presented in Fig.~\ref{fig:specificity} can 
be explained by a combination of priming and  \textit{negative} priming.
Negative priming occurs when a repeated stimulus, which is perceived to be
non-salient, begins to be ignored (\#).

We propose that, as a worker from \textit{task1:food} proceeds through the 
initial tasks, food-related memories and concepts are activated, increasing 
the likelihood that she uses a variety of food-related labels.  This accounts
for the overall increases the number of food-related labels, as well as the 
richness and specificity of her food-related vocabulary.  At the same time,
since all images up to that point contained food, \textit{whether or not}
an image contains food begins to appear non-salient, suppressing the most 
generic references, such as ``food''.  Thus, there is a simultaneous emphasis of the specialized 
references to the prime concept and de-emphasis of generic references.

This view can account for the observation that there are fewer
food-related references overall in \textit{task2:food} compared to 
\textit{task2:cult}.  Workers in \textit{task2:food}, like those in 
\textit{task1:food}, are primed to use
specialized food references and negatively primed against using generic
ones.  Then, when these workers encounter the test tasks, with images 
containing more prominent cultural content, they attend to this novel content, 
decreasing their balance of food references.  
Meanwhile, workers in \textit{task2:cult}, having seen only cultural scenes,
experience novelty from the appearance of food-oriented content in the test 
tasks, and, attending to it, drive the relative difference in the quantity of 
food-related labels further toward the negative in in 
Fig.~\ref{fig:specificity}A.

>>>>

Faced with the scale and complexity of the data revolution, researchers and
industry are increasingly relying on \textit{human computation}.  Microtask
platforms, like Amazon Mechanical Turk, act like a human compute server,
making it possible to access human cognition quickly and inexpensively. 
This has lead to broad adoption across a spectrum of disciplines
in the humanities and computer science, in research into medical imaging
technologies, and in many applications where expert opinion would normally be 
sought.  Here we show that an inherent
feature of microtask platforms, the repetition of similar tasks,
acts as a strong, but as yet unexplored source of bias.  
These \textit{intertask effects}, rival the
severity of overtly framing a task's purpose.  
Using image labeling as a canonical microtask, we found that prior 
tasks influences the vocabulary that a worker uses when labeling 
subsequent images in profound an nuanced ways.  Prior tasks not only 
alter what a worker focuses on, 
but also modulate the richness and specialization of her vocabulary.  
While they can be a source of systematic bias, our
results suggest that intertask effects can be leveraged
to hone worker focus and specificity.  Though we focused on microtask 
platforms due to their emerging prominence, our results apply to  
citizen science, crowdsourcing, and any application involving repetitive 
tasks relying on human judgment.
\end{sciabstract}

\section*{Introduction}
Microtask platforms are online marketplaces on which \textit{requesters} 
post batches of tasks, and \textit{workers} complete them
for remuneration, a sense of purpose, and for fun
\cite{kazai2013analysis,Antin20122925}.  
Typical tasks include tagging and categorizing images 
\cite{6116320,Zhai2012357}, transcribing voice recordings 
\cite{chandler2013breaking,paolacci2010running}
or handwritten notes \cite{Berinsky2012351,Finnerty2013}, and judging the 
relevancy of search results 
\cite{le2010ensuring,grady2010crowdsourcing,alonso2009can,kazai2013analysis}.

While microtask platforms have been widely adopted as a research tool,
computer scientists also hold such platforms as the object of 
research, viewing  microtask platforms as a new kind of computing 
architecture.  In analogy to CPUs (central 
processing units) and GPUs (graphics processing units), 
Davis \textit{et al} invoked the \textit{human processing unit}, or HPU \cite{5543192}.
Meanwhile, researchers have sought to define a basic HPU instruction set, and
libraries for HPU programming are under active development
 \cite{little2010turkit,minder2011crowdlang,minder2012crowdlang,kittur2011crowdforge}.

Unlike CPUs and GPUs, the experiments we present here suggest HPUs
are subject to \textit{hysteresis}, meaning that their output depends
on the history of their inputs.  Given the
increasing prevalence of microtask-based methodologies, other factors 
affecting HPU outputs have received considerable attention.
This includes various task-design parameters such as the level of 
pay \cite{kazai2013analysis}, the training \cite{le2010ensuring} and 
pre-screening \cite{paolacci2010running} of workers, and user-interface 
design factors \cite{Finnerty2013}.  
Researchers have also investigated \textit{framing} effects, 
such as the effects of describing the workflow context 
\cite{Kinnaird2012281}, the purpose of tasks 
\cite{chandler2013breaking}, or altering the problem description
\cite{thibodeau2013natural}.  Results in psychology show that people are 
susceptible to various kinds of priming 
\cite{BJOP1796,No2007,beller1971priming}, and, in particular, 
task-repetition effects \cite{Gass1999549,sohn2001task}.  But to our 
knowledge, no study has investigated the practical ramifications 
of intertask effects on human computation.

We investigated intertask effects using image-labeling tasks on the Amazon 
Mechanical Turk (MTurk) microtask platform.  Image labeling tasks are among
the most common MTurk tasks \cite{chandler2013breaking,Berinsky2012351,Finnerty2013,paolacci2010running}, a trend which will probably continue 
in the long term because of their role in computer vision research 
\cite{5543192}.
In our experiments, workers performed a set of \textit{initial tasks}, 
followed by a set of \textit{test tasks}. Crucially, no distinction was
made between these sets from the perspective of the worker. We varied the 
initial tasks (for example, having workers label images of food or objects), 
and observed the ensuing effects on responses in test tasks.  
As a point of 
comparison, we also included treatments in which we framed the tasks in terms 
of different purposes.

The results showed that intertask effects are stronger than framing effects.  
Framing effects were comparable to intertask effects only when we had the
workers actively reiterate the frame by means of a combo-box input.
Using the wordnet corpus (\#), we found that intertask 
effects can hone a worker's focus, and boost the specialization
and lexical richness of her labels.  Thus, while intertask effects may be a 
serious source of bias, they can also be leveraged to help achieve
expert-level performance in crowdsourcing when properly controlled.

\paragraph{Experimental setup.}
\begin{figure}
	\centering
	\includegraphics[scale=0.7]{figs/images.pdf}
	\caption{
		Examples of images used for various experiments and treatments:
		A) initial tasks for \textit{task1:food}, B) initial tasks for 
		\textit{task1:obj}, C) test tasks for treatments in \textit{exp1}, 
		D) initial tasks for \textit{task2:food}, 
		E) initial tasks for \textit{task2:cult},  
		F) test tasks for \textit{exp2}.  The full set of experimental 
		materials is shown in the supplementary material.
	}

	\label{fig:task}
\end{figure}

We performed two experiments, \textit{exp1} and \textit{exp2}, which consisted
of various pairs of treatments.  The treatments in each pair were subjected to
the same exposure modality, either initial tasks or framing, but differed in
the content of the exposure.  Table \ref{table:treatments}  
summarizes the treatments and introduces a naming convention that we use 
in what follows.
In \textit{exp1}, workers in the treatment pair \textit{task1} initially
labeled 5 images containing either food (treatment \textit{task1:food}), or 
objects, (treatment \textit{task1:obj}), and then performed test tasks, in 
which they 
labeled images containing both food and objects (see Fig.~\ref{fig:task} for 
examples of the images used). In \textit{exp1}, we included two other 
treatment pairs, \textit{frame1} and \textit{echo1}, in which workers did not 
perform initial tasks, but instead were given information which framed the 
purpose of the tasks.  Workers in \textit{frame1} were told that the tasks 
were ``Funded by the laboratory for the visual perception of 
\{ Food and Ingredients $\vert$ Objects and Tools~\}''.  In \textit{echo1},
we sharpened the frame, by indicating that ``The purpose of this study is to 
understand the visual perception of 
\{ Food and Ingredients $\vert$ Objects and Tools~\}'' and then asking the
worker to reiterate the purpose of the task using a combo-box input.

In contrast to \textit{exp1}, which was based on concepts of food and
objects, \textit{exp2} was based food and \textit{culture}.
Since culture is more abstract, one might expect different effects 
in \textit{exp2}.  We again compared
intertask effects (in treatment pair \textit{task2}) to framing effects 
(\textit{frame2}), but with a slight variation in the framing treatments.
In \textit{frame2}, workers were not only exposed to a frame, but also 
performed a set of initial tasks before the test tasks.  Since its purpose 
was still
to investigate framing effects, the initial tasks in \textit{frame2}
were not varied. Thus, \textit{frame2} provides insight into the moderating 
effect that intervening tasks might have on framing effects.

In all treatments, workers were shown images one at a time, and provided
5 labels for each.  From the perspective of the worker, there was no
distinction or interruption between the initial and test tasks.  We used
workers' unique identifiers to ensure that they only participated in a 
one treatment of a single experiment.

@@

	\caption{ \footnotesize{ 
		We use the treatment names shown to refer to specific sections of 
		the experimental data.
	}}
	\label{table:treatments}
\end{table}

\paragraph*{Measuring computational hysteresis.}
While our study builds on research into the psychological mechanisms
of task performance, there is a crucial difference in the question at issue. 
Our goal is to measure the practical ramifications of intertask effects on 
the microtask methodology. Whereas a statistically significant effect, even 
if small, might be sufficient to distinguish one psychological mechanism 
from another (\#), in our study 
\textit{effect size} is central.  Thus our main goal is not merely to reject 
the null hypothesis that initial tasks have no effect
(we do this in the supplementary text).

To couch intertask effects in a practical yet general setting, we imagine
that, after a worker performs a task, her response is used to make some 
kind of yes-or-no decision, $\mathcal{D}$.  For example, this decision 
could be whether or not to flag a certain comment
below a news article as inappropriate; but for now, we leave 
$\mathcal{D}$ unspecified.  Since worker's responses may
vary, repeating the same task might lead to different outcomes for 
$\mathcal{D}$.  We shall define the notion of \textit{bias} in terms of the
effect that prior tasks have on the probability of a positive outcome
from $\mathcal{D}$.

The nature of $\mathcal{D}$ will affect our measurement of bias.
If $\mathcal{D}$ has random output, or if it always outputs ``yes'' 
irrespective of the worker's response, then of course we will 
find zero bias.  Likewise, there is an upper limit to the bias induced on $\mathcal{D}$. In a precise sense, the bias of $\mathcal{D}$ can only be as
great as the bias inherent in the worker's response 
(this is formalized in the supplementary text).  Therefore, to estimate the 
bias induced in workers' outputs, we adopt an 
adversarial approach: we seek to construct the decision that 
is the most susceptible to workers' biases. In the supplementary text, we 
show that this
problem can be formulated as a  machine learning problem, wherein one
constructs a classifier that infers the exposure of a worker from her 
responses in test tasks.  The bias, $\theta$, induced by 
prior exposures, can then be bounded using the classifier's accuracy, 
$\eta$ (\#):
\begin{equation}
	\theta \geq 2\eta - 1
	\label{l1}
\end{equation}

\paragraph{The strength and dynamics of intertask effects.}
For each treatment pair, we trained a naive Bayes classifier (\# in the 
supplementary material we discuss the choice of classifier, the
approach to measuring accuracy, and draw a comparison to results using 
a support vector machine in place of the naive Bayes classifier) to 
distinguish workers from one treatment
(e.g \textit{task1:food}) from those of the conjugate treatment 
(e.g. \textit{task1:obj}).  In \textit{task1} and \textit{task2}, intertask 
effects profoundly biased worker outputs, whereas we could not detect 
a statistically significant bias induced by framing in \textit{frame1} and 
\textit{frame2} (see Fig.~\ref{fig:theta}).  The bias induced
by framing was only of a comparable level to that induced by intertask 
effects in \textit{echo1}.  
Moreover, asking workers to reiterate the
frame, as was done in \textit{echo1}, is probably interpreted as
a direct signal from the requester that the frame should be acted upon.  In that sense, 
\textit{echo1} is perhaps more indicative of the effects of an 
\textit{instruction}.  This shows that
intertask effects profoundly affect the responses of workers, producing an
effect that is stronger than framing, and on par with the combination of 
framing and reiteration.

@@

To observe the dynamics of intertask effects, we trained a naive Bayes
classifier to distinguish workers in \textit{task1} based on responses to
individual test tasks.  We permuted 
the test tasks to eliminate effects arising from differences in the images 
therein (more details are provided in the supplementary text). The first test 
task shows the greatest bias, after which the effects drop off, but 
substantial bias persists through the fifth test task
(see Fig. \ref{fig:theta}B).  Presumably, this decay profile reflects the
fact that workers are also influenced by their exposure to the test tasks, 
which washes out the effects of initial tasks.

@@

Using the wordnet corpus, we analyzed exposure effects in finer detail.
Wordnet encodes semantic information in various ways, including in the form 
of synsets (groups of synonymous words), and hypernym-hyponym relations.
Hypernyms are generalizations (``bread'' is a hypernym of ``pumpernickel''),
while hyponyms are specialisations.

First, we used wordnet to identify nouns depicting foods (\# We 
operationalized food-references by taking all hyponyms of the synsets 
\texttt{food.n.01} and \texttt{food.n.02}, with the addition of ethnic foods 
learned by crawling a recipe website, as described in the supplementary text),
to see whether 
workers from the various food-exposed treatments (\textit{*:food}), 
provided more references to food than their conjugate treatments, overall.
The \textit{exp1:*:food} treatments
did produce more food references, but, unexpectedly, 
the number of food references in \textit{task2:food} was
less than that in \textit{task2:cult}. Also, the effects on the
number of food references was, overall, somewhat smaller than one might 
have expected given the extent of bias observed for \textit{task1}, 
\textit{echo1}, and \textit{task2}.  This means that task and frame exposures
induce other effects which are more subtle than a bulk shift in number 
of references to the primed concept (food).

Next, we looked at the change in the frequency of occurrence of particular 
words.  Surprisingly, the word ``food'' was the \textit{most suppressed} word
for workers in \textit{*:food} across all treatment pairs (Table S1 shows the
top five most promoted and suppressed words for each treatment pair).
Initially, this seems at odds with the fact that (at least in \textit{exp1})
the \textit{*:food} treatments had more food references than their conjugate
treatments.  But taken together, these facts indicate that workers are opting 
to use more specialized food-references in place of ``food''.  
Indeed, when we look at the total number of unique
words (lexical richness) used to refer to food, we found that 
a significantly richer lexicon was used by all \textit{*:food} treatments,
except in \textit{frame2} (see Fig.~\ref{fig:specificity}B).  This effect 
was especially pronounced in \textit{task1} and 
\textit{task2}.

Finally, we directly calculated the degree of specialization of food 
references using wordnet's hypernym-hyponym relations.  To determine the 
relative specialization between the words of two treatments, we computed 
the proportion of words in the first treatment that were hyponyms of the 
those in the 
second, less those in the first that were hypernyms of those in the second 
(this calculation is formalized in the supplementary text).  Except in the
case of \textit{frame2}, workers in \textit{*:food} treatments used 
substantially more specialized terms in reference to food than workers 
from the conjugate treatments.

\paragraph{Positive and negative priming.}
Taken together, the results of the experiments presented here can be 
explained through a combination of positive and negative priming.
Positive priming (usually simply ``priming'') occurs when a prior stimulus 
predisposes a person to give certain responses in an ensuing task, and
is often observed as an increase in the speed or accuracy of a response, or
the ability to recognize objects in a briefer, quieter, or noisier signal (\#).
Negative priming occurs when repeated exposure to a stimulus that is 
perceived to be non-salient begins to be ignored (\#).

Workers exposed to images containing food will be (positively) primed, 
as memories, concepts, and vocabulary related to food are activated.  
However, as the worker labels successive images containing food, the basic 
fact that an image contains food does not appear to be salient, since it 
does not serve to distinguish one image from another.  Thus, the most 
generic reference to the fact that food appears, namely the label ``food'', 
is suppressed, as by negative priming, while more specialized
references are activated.  Whether workers in \textit{*:food} treatments
produce an excess of food references overall depends in part on the balance 
of these factors.  The reduction in food references from workers in 
\textit{task2:food} relative
to those in \textit{task2:cult} may also be related to the novelty that 
workers in 
\textit{task2:cult} experience when they first encounter food in the test 
tasks.

More generally, we are suggesting that, even though workers are not 
instructed to compare the content of tasks in any way, prior tasks form a 
context relative to which workers judge salience.  Thus, a worker's focus 
in repeated tasks tends to be directed away from generic, shared features, 
toward specific and distinguishing ones.

Based on our observations throughout, we submit the following 
recommendations for the consideration of those in the position of designing 
any microtask, citizen science, or crowdsourcing initiative:
\begin{enumerate}
	\item{
		Require active feedback to emphasize important details, 
		as we did in \textit{echo1}.  Similarly, use training tasks rather
		than (only) examples.
	}
	\item{
		Although ``clean'' examples may be useful to provide clarity,
		training tasks should contain the same kinds of noise or otherwise
		distracting features as are expected in the real tasks.
		This exposure will negatively prime the worker against non-salient 
		features.
	}
	\item{
		\textit{Validation tasks}, for which the correct output is known,
		are often used as a quality control measure.
		These should be dually considered as \textit{calibration tasks}.  
		Feedback, for both correct and incorrect responses may further 
		reinforce the desired competency. \label{item:valid}
	}
	\item{
		A special case of item~\ref{item:valid} applies when workers are 
		required to detect rare occurrences.
		For example, in the detection of pre-ictal (pre-seizure) EEG traces, 
		most traces are not negative (\#).  In the absence of positive 
		examples, the target concept may degrade or drift.
		Drawing validation/calibration tasks from the underrepresented class,
		while providing feedback might help sustain the target concept.
	}
	\item{
		For more precise classifications, it may help to begin with
		an initial coarse classification.  For example, in detecting 
		inappropriate comments, workers might initially choose from  
		\texttt{okay}, \texttt{inappropriate}, and \texttt{borderline}.  
		Following the coarse classification, the 
		\texttt{borderline} examples could be batched together.  Workers could
		then calibrate to finer distinctions while coding a more homogeneous 
		set of images.
	}
\end{enumerate}

While these recommendations are informed by the observations we have made 
here, we hope to see them tested directly in future work.
